---
title: "Temporal analysis"
author: "Nicolas Roelandt"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Database connection
library(RPostgreSQL)
library(DBI)
drv <- DBI::dbDriver("PostgreSQL")
#library(sys)

# Data handling
library(sf)
library(dplyr)
library(purrr)

# Graphes
library(ggplot2)
library(scales)

# time / date handling
library(lubridate)
library(hydroTSM)

# Get sun rise time
## install.packages("suncalc")
library(suncalc)
```

```{r con-parameters}
con <- DBI::dbConnect(
drv,
dbname ="noisecapture",
host = "lassopg.ifsttar.fr", #server IP or hostname
port = 5432, #Port on which we ran the proxy
user="noisecapture",
password=Sys.getenv('noisecapture_password') # password stored in .Renviron. Use this to edit it : usethis::edit_r_environ()
)
```




# Data cleaning

The data stored in the database is not useable as is.
Most of the data is not taggued, the accuraccy of the GPS position can vary, the track is too long.
So the data needs to be filtered.

```{r tag-list, include=FALSE}
# Tag list
query <- "SELECT distinct  * FROM noisecapture_tag;"
RPostgreSQL::dbGetQuery(con,statement = query) 
```

This has be done within the database using views.
Please see the [03_create_views.sql](https://github.com/nicolas-roelandt/lasso-data-analysis/blob/main/scripts/03_create_views.sql) script for more details.

In summary, it removes test and indoor records, filter tracks that have a duration between 5 to 900 seconds.
It also remove the tracks where the GPS accuracy was below a 20 meter threshold and that are larger than a 25 square meter area.

This study is focused on France so there is a spatial filtering on its boundaries.

The final data is stored in a view called `france_tracks` that can be called at will.

## Retrieve data
### Track information

```{r join-track-info}
query <- "SELECT  pk_track, record_utc, time_length, pleasantness, noise_level, track_uuid, geog
FROM france_tracks;"

track_info <- sf::st_read(con,query = query)
track_info %>% head()
```

Some records are not in the France metropolitan area so they need to be discarded.

```{r france-metro-filtering}
france_metro <- sf::st_read(dsn=here::here("raw_data/2020_France_metro_WGS84.geojson"))

filtered_track_info <- track_info %>%
  dplyr::filter(sf::st_intersects(., france_metro, sparse = FALSE)) 
```

`r nrow(filtered_track_info)` tracks meets the study criteria (duration, envelop area, etc.).
`r nrow(track_info) - nrow(filtered_track_info)` tracks where not in France's metropolitan area.

### Tag information 

```{r tag-info}
query <- "SELECT ft.pk_track, tag_name FROM france_tracks as ft
INNER JOIN noisecapture_track_tag ntt ON ft.pk_track = ntt.pk_track /* Add track tags*/
INNER JOIN noisecapture_tag ntag ON ntag.pk_tag = ntt.pk_tag /* Add track tags*/;"

tag_info <- RPostgreSQL::dbGetQuery(con,statement = query) %>% dplyr::filter(pk_track %in% filtered_track_info$pk_track)
head(tag_info)
```

Those `r nrow(filtered_track_info)` tracks correspond to `r nrow(tag_info)` tags as a track can have multiple tags.


```{r tag-histo, echo=FALSE}
ggplot(tag_info) +
  aes(x = tag_name) +
  geom_bar(fill = "#112446") +
  coord_flip()+ 
  labs(
    x = "Tag",
    title = "Tag repartition in the subset",
    caption = "Count of each tag"
  ) +
  theme_minimal()
```

# Time repartition
## All year
```{r merge-informations}
all_info <- tag_info %>% 
  dplyr::inner_join(
    filtered_track_info %>% sf::st_drop_geometry()) %>%
  # add local hour
  dplyr::mutate(local_time = lubridate::hour( # extract hour
    lubridate::with_tz( # convert to local time
      lubridate::ymd_hms(record_utc, tz = "UTC"), # convert text to date
      "Europe/Paris")), # target timezone
    season = hydroTSM::time2season(lubridate::date(record_utc), out.fmt = "seasons", type="default")) # compute season

all_info %>% head() %>% knitr::kable()
```


```{r compute-tags-occurrences}
occurences <- all_info %>% dplyr::group_by(tag_name, local_time) %>% dplyr::count(name = "occurences")

occurences %>% head() %>% knitr::kable()
```


```{r compute-tags-hourly-repartition}
tags_hourly_repartition <- occurences %>% 
  left_join(
    occurences %>% dplyr::group_by(local_time) %>% dplyr::summarise(total = sum(occurences)),
    by = "local_time") %>% 
  mutate(percentage = occurences * 100 / total)

tags_hourly_repartition %>% head() %>% knitr::kable()
```


```{r tags-hourly-repartition-graph}
ggplot(tags_hourly_repartition) +
  aes(x = local_time, y = percentage) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(
    x = "Time of the day",
    y = "Percentage",
    title = "Hourly repartition of tags",
    subtitle = "Noicecaptures tags in metropolitan France,
    2017 - 2020"
  ) +
  theme_minimal() +
  facet_wrap(vars(tag_name))

# ggsave("plots/tags_hourly_dynamics.png") 
```

## Seasonal repartition
### Animals

```{r compute-animals-hourly-repartition-season}
seasonal_occurences <- all_info %>% 
  dplyr::group_by(tag_name, local_time, season) %>% 
  dplyr::count(name = "occurences")

animals_seasonal_repartition <- seasonal_occurences %>% 
  dplyr::filter(tag_name == "animals") %>% 
  mutate(seasonal_time = paste0(season,'_',local_time)) %>%
  left_join(
    seasonal_occurences %>%
      dplyr::group_by(local_time, season) %>% 
      dplyr::summarise(total = sum(occurences))%>% 
      dplyr::mutate(seasonal_time = paste0(season,'_',local_time)) %>% select(-season),
    by = "seasonal_time") %>% 
  mutate(percentage = occurences * 100 / total)
  

animals_seasonal_repartition %>% dplyr::select(-local_time.y, -seasonal_time) %>% head() %>% knitr::kable()
```

```{r animals-hourly-repartition-season-graph}
ggplot(animals_seasonal_repartition) +
  aes(x = local_time.x, y = percentage) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(
    x = "Local time",
    y = "Percentage",
    title = "Seasonal occurences of animals sounds"
  ) +
  theme_minimal() +
  facet_wrap(vars(season))

# ggsave("plots/animals_seasonal_dynamic.png") 
```


```{r seasonal-graphs}
season_graphs <- function(tag) {
  seasonal_occurences %>% 
  dplyr::filter(tag_name == tag) %>% 
  mutate(seasonal_time = paste0(season,'_',local_time)) %>%
  left_join(
    seasonal_occurences %>%
      dplyr::group_by(local_time, season) %>% 
      dplyr::summarise(total = sum(occurences))%>% 
      dplyr::mutate(seasonal_time = paste0(season,'_',local_time)) %>% select(-season),
    by = "seasonal_time") %>% 
  mutate(percentage = occurences * 100 / total) %>%
    ggplot() +
  aes(x = local_time.x, y = percentage) +
  geom_point(shape = "circle", size = 1.5, colour = "#112446") +
  labs(
    x = "Local time",
    y = "Percentage",
    title = paste("Seasonal occurences of",tag,"sounds")
  ) +
  theme_minimal() +
  facet_wrap(vars(season))
}

graphs <- purrr::map(unique(seasonal_occurences$tag_name), season_graphs)
```

```{r display-graphs, echo=FALSE, message=FALSE}
graphs
```

# Sound dynamics: sunrise study

```{r compute-sunrise}
get_sunrise <- function(pk_track, date, lat, lon, tz ="UTC") {
  # compute sunrise time from localisation and UTC time
  #return NA if error
  in_pk_track = pk_track
  in_lat = round(lat,5)
  in_lon = round(lon,5)
  in_tz = tz
  
  
  sunrise = tryCatch(suncalc::getSunlightTimes(
    date = lubridate::date(date),
    lat = in_lat,
    lon = in_lon,
    tz   = in_tz
  )$sunrise, error=function(e) NA)

  return(dplyr::tribble(
 ~pk_track, ~sunrise,
 in_pk_track, sunrise)
)
}

# Compute track centroid coordinates
suncalc_prep <- filtered_track_info %>% dplyr::bind_cols(
  filtered_track_info %>% st_centroid() %>% st_coordinates() %>% as_tibble() %>% select(lat = Y, lon = X)
) 

# Compute sunrise hours for each track in a new dataframe
sunrises <- purrr::pmap_dfr(suncalc_prep %>% select(pk_track, date = record_utc, lat, lon) %>% st_drop_geometry, get_sunrise )
```
```{r time-after-sunrise}
# join sunrises to study data
# removes records where sunrise time cannot be compute (track 265404)
time_after_sunrise <- suncalc_prep %>% 
  dplyr::inner_join(sunrises %>% filter(!is.na(sunrise))) %>% #remove NAs to avoid errors later on
  dplyr::mutate(local_time = lubridate::hour( # extract hour
    lubridate::with_tz( # convert to local time
      lubridate::ymd_hms(record_utc, tz = "UTC"), # convert text to date
      "Europe/Paris")),
    local_sunrise = lubridate::hour( # extract hour
    lubridate::with_tz( # convert to local time
      lubridate::ymd_hms(sunrise, tz = "UTC"), # convert text to date
      "Europe/Paris")),
    time_after_sunrise = local_time -local_sunrise
    ) %>% left_join( all_info %>% select(pk_track,tag_name))

time_after_sunrise %>% head() %>% knitr::kable()
```


```{r time-after-sunrise-percentage}
occurences <- time_after_sunrise %>% sf::st_drop_geometry() %>% dplyr::group_by(tag_name, time_after_sunrise) %>% dplyr::count(name = "occurences")

time_after_sunrise_repartition <- occurences %>% 
  left_join(
    occurences %>% dplyr::group_by(time_after_sunrise) %>% dplyr::summarise(total = sum(occurences)),
    by = "time_after_sunrise") %>% 
  mutate(percentage = occurences * 100 / total)

time_after_sunrise_repartition %>% head() %>% knitr::kable()
```


```{r time-after-sunrise-graph}
ggplot(time_after_sunrise_repartition) +
   aes(x = time_after_sunrise, y = percentage) +
   geom_point(shape = "circle", size = 1.5, colour = "#112446") +
   labs(
     x = "Time before or after sunrise",
     y = "Percentage",
     title = "Hourly repartition of tags",
     subtitle = "Noicecaptures tags in metropolitan France,
     2017 - 2020"
   ) +
   theme_minimal() +
   facet_wrap(vars(tag_name))
# ggsave("plots/tag_dynamics_around_sunrise.png") 
```



```{r time-after-sunrise-animals-graph}
ggplot(time_after_sunrise_repartition %>% dplyr::filter(tag_name == "animals")) +
   aes(x = time_after_sunrise, y = percentage) +
   geom_point(shape = "circle", size = 1.5, colour = "#112446") +
   labs(
     x = "Time before or after sunrise",
     y = "Percentage",
     title = "Hourly repartition of tags",
     subtitle = "Noicecaptures tags in metropolitan France,
     2017 - 2020"
   ) +
   geom_smooth(method = "loess", se = FALSE, aes(colour="loess")) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE, aes(colour="Polynomial")) +
  scale_colour_manual(values=c("blue", "red"))+
   theme_minimal()
# ggsave("plots/tag_dynamics_around_sunrise.png") 
```

# Monthly repartition

```{r monthly-repartition}
monthly_occurences <- all_info %>% dplyr::mutate(local_month = lubridate::month( # extract hour
    lubridate::with_tz( # convert to local time
      lubridate::ymd_hms(record_utc, tz = "UTC"), # convert text to date
      "Europe/Paris"))) %>%
  dplyr::group_by(tag_name, local_month, season) %>% 
  dplyr::count(name = "occurences")

tags_monthly_repartition <- monthly_occurences %>%  left_join(
    monthly_occurences %>% dplyr::group_by(local_month) %>% dplyr::summarise(total = sum(occurences)),
    by = "local_month") %>% 
  mutate(percentage = occurences * 100 / total)
```

```{r tags-monthly-repartition-graph}

data_breaks <- data.frame(start = c(0, 3, 6, 9),  # Create data with breaks
                          end = c(3, 6, 9, 12),
                          Seasons = factor(c("Winter", "Spring", "Summer", "Autumn"), 
                                          levels = c("Winter", "Spring", "Summer", "Autumn"),
                          labels = c("Winter", "Spring", "Summer", "Autumn")))
data_breaks                                       # Print data with breaks

ggplot() +
  # Add background colors to plot
  geom_rect(data = data_breaks,
            aes(xmin = start,
                xmax = end,
                ymin = - Inf,
                ymax = Inf,
                fill = Seasons),
            alpha = 0.25) +
  scale_fill_manual(values = c("Winter" = "white",
                               "Spring" = "green",
                               "Summer" = "yellow",
                               "Autumn" = "brown"
                               ))+
  geom_point(data = tags_monthly_repartition, aes(x = local_month, y = percentage),
             shape = "circle", size = 1.5, colour = "#112446") +
  labs(
    x = "Month",
    y = "Percentage",
    title = "Monthly repartition of tags",
    subtitle = "Noicecaptures tags in metropolitan France,
    2017 - 2020"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = scales::pretty_breaks(), labels=c("Jan","Feb","Apr","Jun", "Aug", "Oct", "Dec", "null"), limits = c(1,12))+
  facet_wrap(vars(tag_name))

# ggsave("plots/tags_monthly_dynamics.png") 
```


# TODO


- timezones : https://github.com/Ifsttar/NoiseCapture/tree/master/onomap-geoserver/geoserver/src/test/resources/org/noise_planet/noisecapturegs
  - loaded in database
  - not used because metropolitan France study only




## DONE
- 5s < duration <900s
- GADM28 : remplacer countries ? => NON GADM28 contient des données au niveau communal, il faudrait fusionner 36 000 entités avant d'avoir un résultat acceptable, c'est plus rapide de télécharger le fichier à partir de Natural Earth.
- ajouter url de téléchargement dans table track_info => depuis postgis
```{sql, eval=FALSE}
select record_utc, concat('https://data.noise-planet.org/raw/', substring(user_uuid, 1, 2),'/', substring(user_uuid, 3, 2),'/', substring(user_uuid, 5, 2),'/',user_uuid,'/','track_',track_uuid,'.zip') download  from noisecapture_track nt, noisecapture_user nu where nt.pk_user = nu.pk_user and nt.record_utc > NOW()::date - 1 order by record_utc DESC LIMIT 30;

```
- regarder la répartition par mois également

# Packages citations

```{r}
citation("tidyverse")
```
```{r}
citation("sf")
```

```{r}
citation("hydroTSM")
```


```{r}
citation("suncalc")
```

# Reproductibility
## Data sources
Most of the treatment has been made within the PostGIS database.
The scripts folder contains several scripts to execute to prepare the dataset.

## Session informations

```{r session-info, echo=FALSE}
extract_loaded_package <- function(packages_info) {
  return(packages_info$Package) 
}

xfun::session_info(sapply(sessionInfo()$otherPkgs, extract_loaded_package), dependencies = FALSE)
```

## Database information

```{r pg-version}
# Check database connection and software versions
RPostgreSQL::dbGetQuery(con,statement = paste("SELECT version();")) # should return PostgreSQL 10.15 or higher
```

```{r postgis-version}
RPostgreSQL::dbGetQuery(con,statement = paste("SELECT postgis_full_version();")) # should return PostGIS 2.5 or higher
```

```{r close-connection, echo=FALSE}
RPostgreSQL::dbDisconnect(con)
```


